Parameterization of the responses of subarctic European vegetation to key environmental variables for ozone risk assessment

### 2022-05-30 VegOzone.tex
Editorial changes suggested by Lisa

\begin{abstract}
  The unique vegetation of the subarctic region acclimatized to extremes of cold and midnight sun are likely to be at threat from the combined impacts of climate change and increasing ozone concentrations \chem{[O_3]}. To date, only a few studies assessed the response of subarctic vegetation to variations in climate and air pollution. This study looks to fill this knowledge gap by examining the uncertainty in risk assessments that are using Boreal paramterizations. With no parameterization of stomatal response of subarctic vegetation available, we propose an adjustment of existing Boreal parameterizations based on local climate condition. To this end, we vary temperature response functions according to the climatological temperature distribution within the growing season at our site. Thus, we refer to our approach as localization of stomatal response functions. With these, we apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, key environmental variables, and species-specific physiology. In addition, we evaluate the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and find that 2018 was an anomalously warm and less cloudy year, particularly in spring and early summer. Higher than average \chem{[O_3]} was observed in April/May while frequent episodes of ozone volume mixing ratios (VMRs) exceeding $40\,\unit{ppb}$ occurred in June--August. These episodes are in part attributable to forest fires in the Northern Hemisphere and warmer and sunnier conditions. In contrast, 2019 was an average year. Our modeling study suggests that using standard Boreal parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region for some species. We find that localized parameterizations for subarctic vegetation result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$. For some species, this loss is up to $55\,\unit{\%}$ larger than projected from Boreal parameterizations used in general assessments. Efforts should be targeted towards accurately defining subarctic vegetations' physiological response to essential climate variables.
\end{abstract}

@@

In Section~\ref{sec:dam}, we give a brief summary of the key components and  drivers of the $\mathrm{DO_3SE}$ model that is used to assess the ozone risk. To study the influence of selected parameters on ozone risk in the subarctic, we propose a method to localize temperature response functions based on climate data and to infer start and end of the growing season from satellite observed gross primary productivity (GPP). Results of the ozone risk modelling are presented in Section~\ref{sec:results}. For evaluating these results, we assess the climate anomaly of weather conditions in the 2018/19 GS at our research site in Section~\ref{sec:stats}. In Section~\ref{sec:conc}, we discuss these results concerning prevailing conditions and give an outlook on how future assessments of pollution risk in the subarctic could be improved.

>>>>

\begin{abstract}
  The unique vegetation of the subarctic region, acclimatized to extremes of cold and midnight sun, is likely to be at threat from the combined impacts of climate change and increasing ozone concentrations \chem{[O_3]}. To date, only a few studies have assessed the response of subarctic vegetation to variations in climate and air pollution. This study looks to fill this knowledge gap by examining the uncertainty in risk assessments that result from using Boreal rather than sub-Arctic paramterizations. With no parameterization of stomatal response of subarctic vegetation available, we propose an adjustment of existing Boreal parameterizations based on local climate condition. To this end, we vary the temperature response function according to the climatological temperature distribution during the growing season at our site. We refer to this approach as a 'local climate derived parameterization' of stomatal response functions. With these, we apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, key environmental variables, and species-specific physiology. In addition, we evaluate the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and find that 2018 was an anomalously warm and less cloudy year, particularly in spring and early summer. Higher than average \chem{[O_3]} was observed in April/May while frequent episodes of ozone volume mixing ratios (VMRs) exceeding $40\,\unit{ppb}$ occurred in June--August. These episodes are in part attributable to forest fires in the Northern Hemisphere and warmer and sunnier conditions caused by an extensive blocking event. Such conditions might be more likely in the future under advancing climate change. In contrast, 2019 was an average year. Our modeling study suggests that using standard Boreal parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region for some species. We find that local climate derived parameterizations for subarctic vegetation result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$. For some species, this loss is up to $55\,\unit{\%}$ larger than that projected from Boreal parameterizations used in general assessments. Efforts should be targeted towards accurately defining subarctic vegetations' physiological response to essential climate variables.
\end{abstract}

@@

In Section~\ref{sec:dam}, we give a brief summary of the key components and  drivers of the $\mathrm{DO_3SE}$ model that is used to assess the ozone risk. To study the influence of selected parameters on ozone risk in the subarctic, we propose a method to derive temperature response functions based on local climate and to infer start and end of the growing season from satellite observed gross primary productivity (GPP). Results of the ozone risk modelling are presented in Section~\ref{sec:results}. For evaluating these results, we assess the climate anomaly of weather conditions in the 2018/19 GS at our research site in Section~\ref{sec:stats}. In Section~\ref{sec:conc}, we discuss these results concerning prevailing conditions and give an outlook on how future assessments of pollution risk in the subarctic could be improved.
